State of the Practice for Lattice Boltzmann Method Software

### 2021-12-29 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
updated LBM paper to include Musubi measurements, up to end of paper

Eight packages do not use GitHub or GitLab, so they do not have a measure of
repository stars. Looking at the repository stars of the other 15 packages, we
can observe a slight pattern where packages that have been highly ranked by our
assessment do have more stars than lower ranked packages. The second ranked
package (ESPResSo) has the second most number of stars. The ninth ranked package
(Sailfish) has the most number of stars. Packages designated as lower quality
often do not use GitHub or GitLab, or have few stars, except for HemeLB which
has 22 stars. Our assessment of this package might not be accurate. The number
of stars does not necessarily represent the perceptions of the community, but
for lack of an alternative measure we will use it this way.

The repository watches column contains even less data to compare since two of
the packages (Palabos, waLBerla) use GitLab, which does not include this metric.
The pattern that can be observed with this metric is very similar to that of the
stars metric. The ninth ranked package (Sailfish) has the most number of
watches. The second ranked package (ESPResSo) has the second most number of
watches.

@@

\begin{table}
\begin{center}
		\begin{tabular}{ p{3.5cm}p{2.5cm}p{2cm}p{2cm}p{2cm} }
			\hline
			Name & Our Ranking & Repository Stars & Repository Star Rank & Repository Watches \\
			\hline
			Ludwig&1& 19 &8& 6\\
			ESPResSo &2& 115 &2& 23\\
			Palabos &3& 14 &10& GitLab\\
			OpenLB &4& No Git &No Git & No Git\\
			LUMA&5& 26 &6& 10\\
			pyLBM &6& 57&4& 8\\
			DL\_MESO (LBE) & 7 & No Git &No Git & No Git \\
			waLBerla & 8 & 17 &9& GitLab\\
			Sailfish &9& 154 &1& 39\\
			laboetie &10& 4 &13& 5\\
			TCLB &11& 61 &3& 16\\
			MechSys &12& No Git &No Git & No Git\\
			lettuce &13& 8 &11& 2\\
			ESPResSo++ &14& 31 &5& 13\\
			MP-LABS &15& 7 &12& 2\\			
			SunlightLB & 16& No Git &No Git & No Git\\
			LB3D &17& No Git &No Git & No Git\\			
			LIMBES &18& No Git &No Git & No Git\\
			LB2D-Prime &19& No Git &No Git & No Git\\		
			HemeLB &20& 22 &7& 15\\
			lbmpy&21&  4 &13& 2  \\	
			LB3D-Prime &22& No Git &No Git & No Git\\	
			LatBo.jl &23& 4 &13& 5\\			
			\hline
		\end{tabular}
		\caption{Repository Ranking Metrics} \label{repometrics}
\end{center}
\end{table}

@@

Table~\ref{rankingcomparison} compares the top 10 ranked LBM software packages
with a LBM package ranking made by a domain expert. Five (Palabos, OpenLB, LUMA,
waLBerla, Sailfish) of the top 10 ranked packages in this assessment are also
found in the domain expert's list. Interestingly, these five packages are listed
in the same order on both lists. Looking at the remaining packages, the top two
ranked packages in this state of the practice assessment (Ludwig, ESPResSo) are
not on the domain expert's list. Perhaps the intended applications of these
packages, complex fluids for Ludwig and soft matter research for ESPResSo, do
not align with the research interests of the domain expert, and this is why they
did not make it onto their list. Moving down the list, pyLBM and DL\_MESO(LBE)
did not make the domain expert's top 10 list but were mentioned by the domain
expert. The 10th ranked package, laboetie, was not mentioned by the domain
expert. Looking at the remainder of the domain expert's list, ASL is a general
purpose tool for solving partial differential equations and may have fallen out
of scope of the authoritative lists that were used to identify the initial LBM
software list. The domain expert's fifth ranked package (ch4-project) was on the
initial software list of this assessment, but was removed due to a lack of
documentation. Open FSI is a new project that uses Palabos. It was made public
the year before data was collected for this state of the practice exercise. It
may not have been on authoritative lists due to its young age. Similarly, LIFE
was also made public recently. Finally, the domain expert's 10th ranked package
(HemeLB) was ranked 20th in this state of the practice assessment.

\begin{table}
\begin{center}
		\begin{tabular}{ p{2cm}p{4.5cm}p{3.5cm}}
			\hline
			Rank & This Assessment & Domain Expert\\
			\hline
			1&Ludwig (N/A) &Palabos\\
			2&ESPResSo (N/A) &OpenLB\\
			3&Palabos (1)&LUMA\\
			4&OpenLB (2)&ASL\\
			5&LUMA (3)&ch4-project\\
			6&pyLBM (17)&Open FSI\\
			7&DL\_MESO (LBE) (12)& WaLBerla\\
			8&waLBerla (7)& LIFE\\
			9&Sailfish (9)&Sailfish\\
			10&laboetie (N/A)&HemeLB\\		
			\hline
		\end{tabular}
		\caption{Ranking Comparison} \label{rankingcomparison}
\end{center}
\end{table}

@@

Some pertinent data may not have been specified in public artifacts, or may be
obscure within an artifact or web-page. The use of unit testing and continuous
integration was mentioned in the artifacts of only two (ESPResSo, Ludwig)
packages. However, interviews suggested a more frequent use of both unit testing
and continuous integration in the development processes than what was observed
from the initial survey of the artifacts. For example, OpenLB, pyLBM, and TCLB
use such methods during development despite this not being explicitly clear from
an analysis of the material available online.

>>>>

Nine packages do not use GitHub, so they do not have a measure of
repository stars. Looking at the repository stars of the other 15 packages, we
can observe a slight pattern where packages that have been highly ranked by our
assessment do have more stars than lower ranked packages. The best ranked
package (ESPResSo) has the second most number of stars. The ninth ranked package
(Sailfish) has the most number of stars. Packages designated as lower quality
often do not use GitHub or GitLab, or have few stars. HemeLB only has 12 stars, but its location recently moved from a different GitHub repository. Our assessment of this package might not be accurate. The number of stars does not necessarily represent the perceptions of the community, but for lack of an alternative measure we will use it this way.

The repository watches column contains even less data to compare since two of
the packages (Palabos, waLBerla) use GitLab, which does not include this metric.
The pattern that can be observed with this metric is very similar to that of the
stars metric. The ninth ranked package (Sailfish) has the most number of
watches. The best ranked package (ESPResSo) has the second most number of
watches.

@@

\begin{table}
\begin{center}
		\begin{tabular}{ p{3.5cm}p{2.5cm}p{2cm}p{2cm}p{2cm} }
			\hline
			Name & Our Ranking & Repository Stars & Repository Star Rank & Repository Watches \\
			\hline
			ESPResSo &1& 145 &2& 19\\
			Ludwig&2& 27 &8& 6\\
			Palabos &3& 34 &6& GitLab\\
			OpenLB &4& No Git &No Git & No Git\\
			LUMA&5& 33 &7& 12\\
			pyLBM &6& 95&3& 10\\
			DL\_MESO (LBE) & 7 & No Git &No Git & No Git \\
			Musubi & 8 & No Git &No Git & No Git \\
			Sailfish &9& 186 &1& 41\\
			waLBerla & 10 & 20 &9& GitLab\\
			laboetie &11& 4 &13& 5\\
			TCLB &12& 95 &3& 16\\
			MechSys &13& No Git &No Git & No Git\\
			lettuce &14& 48 &4& 5\\
			ESPResSo++ &15& 35 &5& 12\\
			MP-LABS &16& 12 &11& 2\\			
			SunlightLB & 17& No Git &No Git & No Git\\
			LB3D &18& No Git &No Git & No Git\\			
			LIMBES &19& No Git &No Git & No Git\\
			LB2D-Prime &20& No Git &No Git & No Git\\		
			HemeLB &21& 12 &11& 12\\
			lbmpy&22&  11 &12&  2\\	
			LB3D-Prime &23& No Git &No Git & No Git\\	
			LatBo.jl &24& 17 &10& 8\\			
			\hline
		\end{tabular}
		\caption{Repository Ranking Metrics} \label{repometrics}
\end{center}
\end{table}

@@

Table~\ref{rankingcomparison} compares the top 10 ranked LBM software packages
with a LBM package ranking made by a domain expert. Five (Palabos, OpenLB, LUMA, Sailfish, waLBerla) of the top 10 ranked packages in this assessment are also found in the domain expert's top 10 ranking. Interestingly, the first three of these five packages are listed in the same order on both lists. Looking at the remaining packages, the top two ranked packages in this state of the practice assessment (ESPResSo, Ludwig) are not on the domain expert's list. Perhaps the intended applications of these packages, complex fluids for Ludwig and soft matter research for ESPResSo, do not align with the research interests of the domain expert, and this is why they did not make it onto their list. This reasoning may also apply to eighth ranked Musubi. Moving down the list, pyLBM and DL\_MESO(LBE) did not make the domain expert's top 10 list but were mentioned by the domain expert. Looking at the remainder of the domain expert's list, ASL is a general purpose tool for solving partial differential equations and may have fallen out of scope of the authoritative lists that were used to identify the initial LBM software list. The domain expert's fifth ranked package (ch4-project) was on the initial software list of this assessment, but was removed due to a lack of documentation. Open FSI is a new project that uses Palabos. It was made public the year before data was collected for this state of the practice exercise. It may not have been on authoritative lists due to its young age. Similarly, LIFE was also made public recently. Finally, the domain expert's 10th ranked package (HemeLB) was ranked 21st in this state of the practice assessment.

\begin{table}
\begin{center}
		\begin{tabular}{ p{2cm}p{4.5cm}p{3.5cm}}
			\hline
			Rank & This Assessment & Domain Expert\\
			\hline
			1&ESPResSo (N/A) &Palabos\\
			2&Ludwig (N/A) &OpenLB\\
			3&Palabos (1)&LUMA\\
			4&OpenLB (2)&ASL\\
			5&LUMA (3)&ch4-project\\
			6&pyLBM (16)&Open FSI\\
			7&DL\_MESO (LBE) (11)& WaLBerla\\
			8&Musubi (N/A)& LIFE\\
			9&Sailfish (9)&Sailfish\\
			10&WaLBerla (7)&HemeLB\\		
			\hline
		\end{tabular}
		\caption{Ranking Comparison} \label{rankingcomparison}
\end{center}
\end{table}

@@

Some pertinent data may not have been specified in public artifacts, or may be
obscure within an artifact or web-page. The use of unit testing and continuous
integration was mentioned in the artifacts of only three (ESPResSo, Ludwig, Musubi) packages. However, interviews suggested a more frequent use of both unit testing and continuous integration in the development processes than what was observed from the initial survey of the artifacts. For example, OpenLB, pyLBM, and TCLB use such methods during development despite this not being explicitly clear from an analysis of the material available online.
